---
title: "Lab4"
output: html_document
date: "2023-02-21"
---

```{r setup, include=FALSE}
library(tidyverse)
library(haven)
library(ggthemes)
library(statar) 
library(tidyquant)
library(purrr)
library(sandwich)
library(lmtest)
star <- read_dta("star.dta")
view(star)

```

***Question 1***
It's important to randomly assign both students and teachers to determine any causal effect of class size to prevent selection bias influencing potential given coefficients. It could be that really good teachers could improve test scores, just as really motivated students could improve their test scores as well, confounding the causal effect of class size on test scores. 

```{r Q2}

#Question 2: Average class size comparison between small vs. regular kindergarten classes
#Group data set by whether class size was categorized as small or not then summarize mean class size for each group. 

class_size <- star |> 
  group_by(small) |> 
  summarise(mean = mean(class_size))
class_size

```
The average class size for small classes was roughly 15 students per class; for regular classes, about 22-23 students per class. 

```{r q3a}

#Question 3: Converting different test scores into a standardized unit

#3A: Creating standardized test scores
#creating control variables
control <- star |> 
           filter(small == 0) |> 
           select(read, math, listen, wordskill)

#Math
math_ctrl_mean <- mean(control$math, na.rm = TRUE)
math_ctrl_sd <- sd(control$math, na.rm = TRUE)

math_stdrd <- (star$math - math_ctrl_mean)/math_ctrl_sd

#Reading
read_ctrl_mean <- mean(control$read, na.rm = TRUE)
read_ctrl_sd <- sd(control$read, na.rm = TRUE)

read_stdrd <- (star$read - read_ctrl_mean)/read_ctrl_sd

#Wordskill
wordskill_ctrl_mean <- mean(control$wordskill, na.rm = TRUE)
wordskill_ctrl_sd <- sd(control$wordskill, na.rm = TRUE)

wordskill_stdrd <- (star$wordskill - wordskill_ctrl_mean)/wordskill_ctrl_sd

#Listening 
listen_ctrl_mean <- mean(control$listen, na.rm = TRUE)
listen_ctrl_sd <- sd(control$listen, na.rm = TRUE)

listen_stdrd <- (star$listen - listen_ctrl_mean)/listen_ctrl_sd

#Reporting summary statistics for four standardized test score variables

summary(math_stdrd)
sd(math_stdrd)

summary(read_stdrd)
sd(read_stdrd)

summary(wordskill_stdrd)
sd(wordskill_stdrd)

summary(listen_stdrd)
sd(listen_stdrd)

```

```{r q3b}

#Creating standardized test score across all scores

sat_index <- (math_stdrd + read_stdrd + wordskill_stdrd + listen_stdrd)/4
summary(sat_index)
```

```{r q3c}

#3C: Histogram of sat_index for small and regular kindergarten classes

histosat <- star |> 
ggplot(aes(x = sat_index,
          fill = factor(small, labels=c("Small", "Regular")),
          y = ..density..)) +
          geom_histogram(alpha = 0.5, position = "identity") +
          labs(x = "Standardized Test Score", 
               y = "Frequency",
               fill = "Class Size",
               title = "Standardized SAT Scores for Small vs. Regular Classrooms") 
histosat

ggsave("histostat.png")
```
There is some overlap between the standardized test scores for small and regular classrooms. 


```{r q4}

#4A: Aggregating the dataset by teacher information

by_class <- star |> 
            group_by(teacher_id, school_id, small, teacher_masters, teacher_white, teacher_black,    teacher_experience)

classes <- by_class |> 
           summarise(sat_index = mean(sat_index),
                     class_size = mean(class_size))

#reporting summary variables 
summarysmall <- classes |> 
                filter(small == 1) |> 
                summary() 
summarysmall

summaryreg <- classes |> 
                filter(small == 0) |> 
                summary() 
summaryreg

```

```{r q4b}

#4B: Creating linear model for teacher experience and class size, constructing 95% confidence interval
mod1 <- lm(teacher_experience ~ small, data = classes)
summary(mod1)

#95% Confidence interval
mod1coef <- coeftest(mod1, vcov = vcovHC(mod1, type="HC1"))
mod1se <- mod1coef[4]
estdiff <- mod1coef[2]
lb_cf1 <- estdiff - 1.96*mod1se
ub_cf1 <- estdiff + 1.96*mod1se

lb_cf1 #and 
ub_cf1

```

The 95 % confidence interval for the estimated difference in teacher experience based on class size is between a -1.71 decrease in years of teacher experience and an .89 increase in years of experience. The regression coefficient is not statistically significant. 

```{r q4c}

#4C: Teacher Masters
mod2 <- lm(teacher_masters ~ small, data = classes)
summary(mod2)

#95% Confidence interval
mod2coef <- coeftest(mod2, vcov = vcovHC(mod2, type="HC1"))
mod2se <- mod2coef[4]
estdiff <- mod2coef[2]
lb_cf2 <- estdiff - 1.96*mod2se
ub_cf2 <- estdiff + 1.96*mod2se

lb_cf2 #and 
ub_cf2

```

The 95 % confidence interval for the estimated difference in a teacher having a master's degree based on class size is between a -0.14 decrease in the likelihood of a teacher having a master's and a 0.06 increase in the likelihood of a given teacher having a master's. The regression coefficient is not statistically significant. 

```{r q4c}

#4c: Teacher white
mod3 <- lm(teacher_white ~ small, data = classes)
summary(mod3)

#95% Confidence interval
mod3coef <- coeftest(mod3, vcov = vcovHC(mod3, type="HC1"))
mod3se <- mod3coef[4]
estdiff <- mod3coef[2]
lb_cf3 <- estdiff - 1.96*mod3se
ub_cf3 <- estdiff + 1.96*mod3se

lb_cf3 #and 
ub_cf3

```

The 95 % confidence interval for the estimated difference in a teacher being white based on class size is between a -0.03 decrease in the likelihood of a teacher being white and a 0.12 increase. The regression coefficient is not statistically significant. 

```{r q4c}

#4C: Teacher Black
mod4 <- lm(teacher_black ~ small, data = classes)
summary(mod4)

#95% Confidence interval
mod4coef <- coeftest(mod4, vcov = vcovHC(mod4, type="HC1"))
mod4se <- mod4coef[4]
estdiff <- mod4coef[2]
lb_cf4 <- estdiff - 1.96*mod4se
ub_cf4 <- estdiff + 1.96*mod4se

lb_cf4 
ub_cf4

```

The 95 % confidence interval for the estimated difference in a teacher being Black based on class size is between a -0.11 decrease in the likelihood of a teacher being white and a 0.04 increase. The regression coefficient is not statistically significant. 

***Question 4D***
As reported above, none of the regression coefficients in the linear models exploring the relationship between each of the teacher characteristics and class size were statistically significant. The coefficient magnitudes are quite small compared to the intercept, suggesting that they are not practically significant. This suggests that the randomization was successful in balancing teacher characteristics. 

```{r q5}
#Building multilinear regression model using school fixed effects

mod5 <- lm(sat_index ~ small + factor(school_id), data = classes)
summary(mod5)

#95% confidence interval
mod5coef <- coeftest(mod5, vcov = vcovHC(mod5, type="HC1"))
mod5se <- mod5coef[2,2]
estdiff <- mod5coef[2]
lb_cf5 <- estdiff - 1.96*mod5se
ub_cf5 <- estdiff + 1.96*mod5se

#Bar chart showing the difference in standardized test score between small and regular classes
control_bar <- mean(classes$sat_index[classes$small == 0])
treat_bar <- control_bar + estdiff

#create a dataframe
data_for_graph <- data_frame(c(control_bar, treat_bar),
                             c(NA, mod5se),
                             c("Control", "Treatment"),
                             c(NA, lb_cf5),
                             c(NA, ub_cf5))

names(data_for_graph)[1] <- "SAT_Index"
names(data_for_graph)[2] <- "SE"
names(data_for_graph)[3] <- "Groups"
names(data_for_graph)[4] <- "Lower_bound"
names(data_for_graph)[5] <- "Upper_bound"

#create graph
bar <- data_for_graph |> 
       ggplot(aes(x = Groups, y = SAT_Index, fill = Groups)) +
       geom_bar(stat = "identity", show.legend = FALSE, width = .6) +
       geom_errorbar(aes(ymin = Lower_bound, ymax = Upper_bound), width=.1, size = 0.7,      color="black") +
      labs(y = "SAT Index Score",
           title = "Estimated Effect of STAR Treatment on Test Scores") +
      scale_fill_manual(values = c("red", "blue")) 

ggsave("bar.png")



```



